The effect of habitat structure on microbial community function and composition

### 2019-11-07 SI_code_revised.Rmd
Last minor changes

#Bacteria
bac<-as.data.frame(t(read.csv(file=c("bacteria.csv"), skip = 4)[,-1]))
colnames(bac)<-t(read.csv(file=c("bacteria.csv"), skip = 4)[,1])

@@

@@

```{r vegan analysis, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#Bacteria
bac.norm<-decostand(bac, method=c("normalize"))
bac.dist<-vegdist(bac.norm, method = "bray")
bac.ad<-adonis(bac.dist~Substrate+Structure/Substrate, bac_env, strata = bac_env$Day)

#Fungi
fungi.norm<-decostand(fungi, method=c("normalize"))
fungi.dist<-vegdist(fungi.norm, method = "bray")
fungi.ad<-adonis(fungi.dist~Substrate+Structure/Substrate, fungi_env, strata = fungi_env$Day)

@@

kable(t.s1.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$'), row.names=T, caption = "_**Table S3:** Monod model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

kable(t.s2.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$', 'k'), row.names=T, caption = "_**Table S5:** Monod plus model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

kable(t.s3.1, digits = c(2, 2, 3, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', '$\\m_{R}$', '$\\f_{0}$', 'k', '$\\Y_{S}$'), row.names=T, caption = "_**Table S7:** Two pool model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

kable(t.s4.1, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$', '$\\f_{Proteins}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S9:** Monod model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

kable(t.s4.2, digits = c(0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$', '$\\f_{Proteins}$'), row.names=F, caption = "_**Table S11:** Monod model parameters calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

kable(t.s4.3, digits = c(0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S13:** Monod model parameters calibrated against respiration rate and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

kable(t.s5.1, digits = c(0, 2, 2, 2, 3, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$', '$\\k_{MB}$', '$\\f_{Proteins}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S15:** Monod plus model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

kable(t.s5.2, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$', '$\\k_{MB}$', '$\\f_{Proteins}$'), row.names=F, caption = "_**Table S17:** Monod plus model parameters calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

kable(t.s5.3, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$', '$\\k_{MB}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S19:** Monod plus model parameters calibrated against respiration rate and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

####Two pool model
Not all model parameters vary significantly acros treatments. Parameters that doesn't show any variability across treatments are fixed.
```{r Two pool model calibration, eval=F, echo=T}
#Loading the function 
source("../Rscripts/two_pool_all.R")
source("../Rscripts/two_pool_all_fix.R")
#across all treatments
two_pool_all1<-two_pool_all(data=d, FACT = 4)
###########################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
###########################################################
#for different structures
two_pool_all2<-two_pool_all(data=d, FACT = 2)
two_pool_all2_fix<-two_pool_all_fix(data=d, FACT = 2)
############################################################################
#in the following code, the Ys and k parameters are fixed across treatments
#becuase its variability across treatments is minimal (data not shown)
source("./Rscripts/two_pool_all_fixYuk.R")
two_pool_all_fixYuk2<-two_pool_all_fixYuk(data=d, FACT = 2)
#Then, parameter f is fixed as well
source("./Rscripts/two_pool_all_fixYukf.R")
two_pool_all_fixYukf2<-two_pool_all_fixYukf(data=d, FACT = 2)
#Then, parameter Vmax is fixed as well
source("./Rscripts/two_pool_all_fixYukfVmax.R")
two_pool_all_fixYukfVmax2<-two_pool_all_fixYukfVmax(data=d, FACT = 2)
#At the end, parameter KM is fixed
source("./Rscripts/two_pool_all_fixYukfVmaxKm.R")
two_pool_all_fixYukfVmaxKm2<-two_pool_all_fixYukfVmaxKm(data=d, FACT = 2)
############################################################################
#for different substrates
two_pool_all3<-two_pool_all(data=d, FACT = 1)
two_pool_all3_fix<-two_pool_all_fix(data=d, FACT = 1)
#for each separately 
two_pool_all4<-two_pool_all(data=d, FACT = 3)
two_pool_all4_fix<-two_pool_all_fix(data=d, FACT = 3)
#in the following code, the Ys, k,f and Vmax parameters are fixed across treatments
two_pool_all_fixYukfVmax4<-two_pool_all_fixYukfVmax(data=d, FACT = 3)
###########################################################
stopImplicitCluster()
###########################################################
two_pool_all1$goodness
two_pool_all_fixYukfVmaxKm2$goodness
two_pool_all3_fix$goodness
two_pool_all_fixYukfVmax4$goodness
```

@@

#results of the previous iteration (not run)

@@

#                             Km = 6.3136,
#                             two_pool_all_fixYukfVmaxKm2[[1]]$pars[1],

@@

#                             two_pool_all_fixYukfVmaxKm2[[1]]$pars[c(2:4)],

@@

#                             Km = 6.3136,
#                             two_pool_all_fixYukfVmaxKm2[[2]]$pars[1],

@@

#                             two_pool_all_fixYukfVmaxKm2[[2]]$pars[c(2:4)],

@@

#                             Km = 6.3136,
#                             two_pool_all_fixYukfVmaxKm2[[3]]$pars[1],

@@

#                             two_pool_all_fixYukfVmaxKm2[[3]]$pars[c(2:4)],

@@

#                                  two_pool_all_fixYukfVmaxKm2$goodness,

@@

kable(t.s6.1, digits = c(0, 2, 2, 3, 3, 3, 2, 3, 3, 3, 2), align = 'c', "html", booktabs=T, col.names = c(" ","$\\V_{MAX}$", '$\\K_{M}$', '$\\m_{R}$', '$\\f_{0}$', 'k', '$\\Y_{S}$', '$\\R_{Proteins}$', '$\\S_{Proteins}$', '$\\S_{DNA}$', '$\\RS_{0}$'), row.names=F, caption = "_**Table S21:** Two pool model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

$~$
$~$
$~$
$~$
$~$
$~$
```{r correlation figure, echo=FALSE, fig.height=5, fig.width=10, fig.cap="Fig. S3: Effect of structure of bacterial and fungal part of microbial community (expressed as multivariate mean distance between groups; see Matrial and Methods section) on overall microbial community maintnance respiration rate ($m_{R}$ - (a)) and protein concentration in Structures pool ($S_{Proteins}$ - (b)). The value of both parameters are expressed as a mean difference between groups in order to allow unbiased correlation with microbial community structure metric (see Matrial and Methods section)."}

dpars<-t.s6.1[c(2:4), -1]
dpars$Structure<-c("BROTH", "GLASS", "WOOL")

>>>>

#Bacteria
bac<-as.data.frame(t(read.csv(file=c("bacteria.csv"), skip = 4)[,-1]))
colnames(bac)<-t(read.csv(file=c("bacteria.csv"), skip = 4)[,1])
##remove OTUs with all zeros
bac2<-bac[, colSums(bac!=0)>0]

@@

##remove OTUs with all zeros
fungi2<-fungi[, colSums(fungi!=0)>0]

@@

```{r vegan analysis, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#Bacteria
bac.norm<-decostand(bac2, method=c("normalize"))
bac.dist<-vegdist(bac.norm, method = "bray")
bac.ad<-adonis(bac.dist~Substrate+Structure/Substrate, bac_env, strata = bac_env$Day)

#Fungi
fungi.norm<-decostand(fungi2, method=c("normalize"))
fungi.dist<-vegdist(fungi.norm, method = "bray")
fungi.ad<-adonis(fungi.dist~Substrate+Structure/Substrate, fungi_env, strata = fungi_env$Day)

@@

kable(t.s1.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('v', 'k', '$\\Y_{CUE}$', 'm'), row.names=T, caption = "_**Table S3:** Monod model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

kable(t.s2.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('v', 'k', '$\\Y_{CUE}$', 'h', 'm'), row.names=T, caption = "_**Table S5:** Monod plus model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

kable(t.s3.1, digits = c(2, 2, 3, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('v', 'k', 'h', 'f', 'm', 'Y'), row.names=T, caption = "_**Table S7:** Two pool model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

kable(t.s4.1, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{Proteins}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S9:** Monod model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

kable(t.s4.2, digits = c(0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{Proteins}$'), row.names=F, caption = "_**Table S11:** Monod model parameters calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

kable(t.s4.3, digits = c(0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S13:** Monod model parameters calibrated against respiration rate and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

kable(t.s5.1, digits = c(0, 2, 2, 2, 3, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'h', 'm', '$\\f_{Proteins}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S15:** Monod plus model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

kable(t.s5.2, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'h', 'm', '$\\f_{Proteins}$'), row.names=F, caption = "_**Table S17:** Monod plus model parameters calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

kable(t.s5.3, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'h', 'm', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S19:** Monod plus model parameters calibrated against respiration rate and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

####Two pool model
Not all model parameters vary significantly across treatments. Parameters that doesn't show any variability across treatments are fixed.
```{r Two pool model calibration, eval=F, echo=T}
#Loading the function 
source("../Rscripts/two_pool_all.R")
source("../Rscripts/two_pool_all_fix.R")
#across all treatments
two_pool_all1<-two_pool_all(data=d, FACT = 4)
###########################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
###########################################################
#for different structures
two_pool_all2<-two_pool_all(data=d, FACT = 2)
two_pool_all2_fix<-two_pool_all_fix(data=d, FACT = 2)
############################################################################
#in the following code, the Ys and k parameters are fixed across treatments
#becuase its variability across treatments is minimal (data not shown)
source("./Rscripts/two_pool_all_fixYuk.R")
two_pool_all_fixYuk2<-two_pool_all_fixYuk(data=d, FACT = 2)
#Then, parameter f is fixed as well
source("./Rscripts/two_pool_all_fixYukf.R")
two_pool_all_fixYukf2<-two_pool_all_fixYukf(data=d, FACT = 2)
#Then, parameter Vmax is fixed as well
source("./Rscripts/two_pool_all_fixYukfVmax.R")
two_pool_all_fixYukfVmax2<-two_pool_all_fixYukfVmax(data=d, FACT = 2)
############################################################################
#for different substrates
two_pool_all3<-two_pool_all(data=d, FACT = 1)
two_pool_all3_fix<-two_pool_all_fix(data=d, FACT = 1)
#for each separately 
two_pool_all4<-two_pool_all(data=d, FACT = 3)
two_pool_all4_fix<-two_pool_all_fix(data=d, FACT = 3)
#in the following code, the Ys, k,f and Vmax parameters are fixed across treatments
two_pool_all_fixYukfVmax4<-two_pool_all_fixYukfVmax(data=d, FACT = 3)
###########################################################
stopImplicitCluster()
###########################################################
two_pool_all1$goodness
two_pool_all_fixYukfVmax2$goodness
two_pool_all3_fix$goodness
two_pool_all_fixYukfVmax4$goodness
```

@@

# #results of the previous iteration (not run)

@@

#                            two_pool_all_fixYukfVmax2[[1]]$pars[1],
#                            two_pool_all_fixYukfVmax2[[1]]$pars[2],

@@

#                            two_pool_all_fixYukfVmax2[[1]]$pars[c(3:5)],

@@

#                            two_pool_all_fixYukfVmax2[[2]]$pars[1],
#                            two_pool_all_fixYukfVmax2[[2]]$pars[2],

@@

#                            two_pool_all_fixYukfVmax2[[2]]$pars[c(3:5)],

@@

#                            two_pool_all_fixYukfVmax2[[3]]$pars[1],
#                            two_pool_all_fixYukfVmax2[[3]]$pars[2],

@@

#                            two_pool_all_fixYukfVmax2[[3]]$pars[c(3:5)],

@@

#                                 two_pool_all_fixYukfVmax2$goodness,

@@

kable(t.s6.1, digits = c(0, 2, 2, 3, 3, 3, 2, 3, 3, 3, 2), align = 'c', "html", booktabs=T, col.names = c(" ","v", 'k', 'h', 'f', 'm', 'Y', '$\\B_{R-Proteins}$', '$\\B_{S-Proteins}$', '$\\B_{S-DNA}$', '$\\RS_{0}$'), row.names=F, caption = "_**Table S21:** Two pool model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

$~$
$~$
$~$
$~$
$~$
$~$
```{r correlation figure, echo=FALSE, fig.height=12, fig.width=16, fig.cap="Fig. S3: Relative abundance of bacterial (a) and fungal (b) species in three experimental treatments differing in presence/absence of physical structure. Species with relative abundance above 0.5% are plotted. Their abundance is visualized by colour scale. Effect of bacterial and fungal composition of microbial community (expressed as multivariate mean distance between groups; see Material and Methods section) on overall microbial community affinity to substrate (i.e. parameter k - (a)) and abundance of proteins in microbial biomass (i.e. sum of protein abundance in biomass reserve and struscture - (b)). The values on y-axes are expressed as a mean difference between groups in order to allow unbiased correlation with microbial community structure metric (see Matrial and Methods section)."}
#Heatmap
bac.tax<-as.data.frame(read.csv(file=c("./Community_analysis/OTU_bacteria.txt"), header=T,
                             sep="\t"))
bac.tax<-bac.tax[, c(1, 175)]
bac.tax$X.OTU.ID<-gsub("_","", bac.tax$X.OTU.ID)

#Add taxonomy to bac.norm dataset
bac.norm_tax<-bac
#Transform the dataset
bac.norm_taxt<-as.data.frame(t(bac.norm_tax))
taxlab<-character()
for(i in 1:nrow(bac.norm_taxt)){
  nd<-which(bac.tax$X.OTU.ID==rownames(bac.norm_taxt)[i])
  taxlab<-append(taxlab, as.character(bac.tax$taxonomy[nd]))
}

bac.norm_taxt$taxonomy<-taxlab

#Extract kingdom
pattern<-"k__(.+?),"
regmatches(bac.norm_taxt$taxonomy[1], regexec(pattern, bac.norm_taxt$taxonomy[1]))[[1]][2]
kingdoms<-character()
for(i in 1:nrow(bac.norm_taxt)){
  kingdoms<-append(kingdoms, 
                   regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
}
kingdoms[which(kingdoms=="?")]<-c("Unknown")
bac.norm_taxt$Kingdom<-kingdoms

#Extract phylum
pattern<-"p__(.+?),"

phylum<-character()
for(i in 1:nrow(bac.norm_taxt)){
  phylum<-append(phylum, 
                   regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
}
phylum[which(phylum=="?")]<-c("Unknown")
bac.norm_taxt$Phylum<-phylum

#Extract class
pattern<-"c__(.+?),"

cs<-character()
for(i in 1:nrow(bac.norm_taxt)){
  cs<-append(cs, 
                 regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
}
cs[which(cs=="?")]<-c("Unknown")
bac.norm_taxt$Class<-cs

#Extract order
pattern<-"o__(.+?),"

os<-character()
for(i in 1:nrow(bac.norm_taxt)){
  os<-append(os, 
             regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
}
os[which(os=="?")]<-c("Unknown")
bac.norm_taxt$Order<-os

#Extract family
pattern<-"f__(.+?),"

fs<-character()
for(i in 1:nrow(bac.norm_taxt)){
  fs<-append(fs, 
             regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
}

fs[which(fs=="?")]<-c("Unknown")
bac.norm_taxt$Family<-fs

#Extract genus
pattern<-"g__(.+?),"

gs<-character()
for(i in 1:nrow(bac.norm_taxt)){
  gs<-append(gs, 
             regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
}
gs[which(gs=="?")]<-c("Unknown")
bac.norm_taxt$Genus<-gs

#Extract species
pattern<-"s__(.+?),"

ss<-character()
for(i in 1:nrow(bac.norm_taxt)){
  ss<-append(ss, 
             regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
}
ss[which(is.na(ss))]<-c("Unknown")
bac.norm_taxt$Specie<-ss

#Plots
tax_env<-bac_env
btaxt<-bac.norm_taxt[, 1:71]
tax_env$IDs<-colnames(btaxt)

##Genus
btaxtg<-btaxt
btaxtg$Genus<-bac.norm_taxt$Genus

Btaxtg<-melt(btaxtg, id.vars = c("Genus"))
colnames(Btaxtg)<-c("Genus", "IDs","Abundance")

Btaxtg<-merge(Btaxtg, tax_env, by.x = "IDs", by.y = "IDs", all.x = TRUE)
Btaxtg$Total<-numeric(length = nrow(Btaxtg))
for(i in unique(Btaxtg$IDs)){
  Btaxtg[Btaxtg$IDs==i, "Total"]<-sum(as.numeric(Btaxtg[Btaxtg$IDs==i, "Abundance"]))
}

bac_gen_f<-Btaxtg %>% group_by(Genus, Structure, Substrate) %>% 
  summarize(Abundance=mean(Abundance/Total, na.rm=T))

#Only genus with abundance above 0.05%
heat1<-ggplot(bac_gen_f[bac_gen_f$Abundance>0.0005, ], aes(Structure, Genus)) +
  geom_tile(aes(fill = log10(Abundance*100))) +
  facet_grid(~ Substrate) +
  scale_fill_gradientn(colours = c("white","darkorange2","grey60"),
                       breaks = c(-2, -1, 0, 1, 2),
                       labels=c("0.01", "0.1", "0", "10", "100"),
                       name = "Relative abundance (%)",
                       guide = guide_colourbar(direction = "horizontal", title.position = "bottom",
                                               title.hjust = 0.5, barwidth = 12)) +
  scale_x_discrete(expand = c(0,0)) + 
  scale_y_discrete(expand = c(0,0)) +  
  xlab("Structure") + ylab("")+
  ggtitle("a)")+
  theme_min + theme(legend.position = "bottom",
                    axis.text = element_text(size=14),
                    strip.text = element_text(size=14))

##Fungi
f.tax<-as.data.frame(read.csv(file=c("./Community_analysis/OTU_fungi.txt"), header=T,
                             sep="\t"))
#Add taxonomy to fungi dataset
f.norm_tax<-fungi
#Transform the dataset
f.norm_taxt<-as.data.frame(t(f.norm_tax))
ftaxlab<-character()
for(i in 1:nrow(f.norm_taxt)){
  nd<-which(f.tax$X.OTU.ID==rownames(f.norm_taxt)[i])
  ftaxlab<-append(ftaxlab, as.character(f.tax$taxonomy[nd]))
}

f.norm_taxt$taxonomy<-ftaxlab

#Extract kingdom
pattern<-"k__(.+?),"
regmatches(f.norm_taxt$taxonomy[1], regexec(pattern, f.norm_taxt$taxonomy[1]))[[1]][2]
kingdoms<-character()
for(i in 1:nrow(f.norm_taxt)){
  kingdoms<-append(kingdoms, 
                   regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
}
kingdoms[which(kingdoms=="?")]<-c("Unknown")
f.norm_taxt$Kingdom<-kingdoms

#Extract phylum
pattern<-"p__(.+?),"

phylum<-character()
for(i in 1:nrow(f.norm_taxt)){
  phylum<-append(phylum, 
                 regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
}
phylum[which(phylum=="?")]<-c("Unknown")
f.norm_taxt$Phylum<-phylum

#Extract class
pattern<-"c__(.+?),"

cs<-character()
for(i in 1:nrow(f.norm_taxt)){
  cs<-append(cs, 
             regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
}
cs[which(cs=="?")]<-c("Unknown")
f.norm_taxt$Class<-cs

#Extract order
pattern<-"o__(.+?),"

os<-character()
for(i in 1:nrow(f.norm_taxt)){
  os<-append(os, 
             regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
}
os[which(os=="?")]<-c("Unknown")
f.norm_taxt$Order<-os

#Extract family
pattern<-"f__(.+?),"

fs<-character()
for(i in 1:nrow(f.norm_taxt)){
  fs<-append(fs, 
             regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
}

fs[which(fs=="?")]<-c("Unknown")
f.norm_taxt$Family<-fs

#Extract genus
pattern<-"g__(.+?),"

gs<-character()
for(i in 1:nrow(f.norm_taxt)){
  gs<-append(gs, 
             regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
}
gs[which(gs=="?")]<-c("Unknown")
f.norm_taxt$Genus<-gs

#Extract species
pattern<-"s__(.+?),"

ss<-character()
for(i in 1:nrow(f.norm_taxt)){
  ss<-append(ss, 
             regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
}
ss[which(is.na(ss))]<-c("Unknown")
f.norm_taxt$Specie<-ss

#Plots
ftax_env<-fungi_env
ftaxt<-f.norm_taxt[, 1:64]
ftax_env$IDs<-colnames(ftaxt)

##Genus
ftaxtg<-ftaxt
ftaxtg$Genus<-f.norm_taxt$Genus

Ftaxtg<-melt(ftaxtg, id.vars = c("Genus"))
colnames(Ftaxtg)<-c("Genus", "IDs","Abundance")

Ftaxtg<-merge(Ftaxtg, ftax_env, by.x = "IDs", by.y = "IDs", all.x = TRUE)
Ftaxtg$Total<-numeric(length = nrow(Ftaxtg))
for(i in unique(Ftaxtg$IDs)){
  Ftaxtg[Ftaxtg$IDs==i, "Total"]<-sum(as.numeric(Ftaxtg[Ftaxtg$IDs==i, "Abundance"]))
}

f_gen_f<-Ftaxtg %>% group_by(Genus, Structure, Substrate) %>% 
  summarize(Abundance=mean(Abundance/Total, na.rm=T))
